Gravitational wave sensitivity curves

### 2014-08-01 Detectors.tex
RHC and CJM alterations

The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600. LIGO, Virgo and GEO600 have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo). They are currently being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV). LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana. There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}. The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}, this is not included in Table \ref{table:t} \cplb{Why not?}. The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.

>>>>

The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600. LIGO, Virgo and GEO600 have completed science runs as first-generation detectors. As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively. The upgraded, second-generation versions are referred to as advanced LIGO (aLIGO) and advanced Virgo (AdV) respectively. LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana. There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}. The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}. The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.

### 2014-08-01 Typesofsource.tex
RHC and CJM alterations

The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation. Consider a source signal with approximately constant (root-mean-square) amplitude $h_0$ and central frequency $f'$. In this case,
\begin{eqnarray}
h(t) &=& \sqrt{2} h_0 \cos\left[\rmi \phi(t)\right] \; ; \\
\tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
\label{eq:fourier-h}
\end{eqnarray}
Without loss of generality, we can assume an initial phase of zero, such that $\phi(0) = 0$. The largest contribution to the integral comes from where the argument of the exponentials is approximately zero. For the first term, this occurs when $f = f'$, then the term in brackets is
\begin{equation}
\left[\frac{\phi(t)}{2\pi t}-f\right]_{f\,=\,f'} = f' + \dot{f'}t + \mathcal{O}\left(t^2\right) - f' = \dot{f'}t + \mathcal{O}\left(t^2\right).
\end{equation}
The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected. Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) gives
\begin{eqnarray} \label{eq:FTofEMRI}
\tilde{h}(f) &\simeq& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp\left(2\pi\rmi \dot{f'} t^{2}\right) + \exp\left(-2\pi\rmi \dot{f'} t^{2}\right) \nonumber \\
&\simeq& \frac{h_0}{\sqrt{2\dot{f'}}}.
\end{eqnarray}
From (\ref{eq:strain-hc}) and (\ref{eq:FTofEMRI}), the characteristic strain for inspiralling sources is given by \citep{FinnThorne}
\begin{equation}\label{eq:insphc}
h_\mathrm{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h_0 \;.
\end{equation}
Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals. Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.

@@

An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations:
\begin{equation}
h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
\end{equation}
If $\tilde{h}(f)$ is constant across the bandwidth $\Delta f$, this is related to the characteristic strain via \cplb{What about the detector response?}. In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.

>>>>

The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation. Consider a source signal with approximately constant (root-mean-square) amplitude $h_0$ and central frequency $f'$. In this case,
\begin{eqnarray}
h(t) &=& \sqrt{2} h_0 \cos\left[\phi(t)\right] \; , \\
\tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
\label{eq:fourier-h}
\end{eqnarray}
Without loss of generality, we can assume an initial phase of zero, such that $\phi(0) = 0$. The largest contribution to the integral comes from where the argument of the exponentials is approximately zero. For the first term, this occurs when $f = f'$, then the term in brackets is
\begin{equation}
\left[\frac{\phi(t)}{2\pi t}-f\right]_{f\,=\,f'} = f' + \dot{f'}t + \mathcal{O}\left(t^2\right) - f' = \dot{f'}t + \mathcal{O}\left(t^2\right).
\end{equation}
The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected. Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) and then evaluating the Gaussian integrals gives
\begin{eqnarray} \label{eq:FTofEMRI}
\tilde{h}(f) &\simeq& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp\left(2\pi\rmi \dot{f'} t^{2}\right) + \exp\left(-2\pi\rmi \dot{f'} t^{2}\right) \nonumber \\
&\simeq& \frac{h_0}{\sqrt{2\dot{f'}}}.
\end{eqnarray}
From (\ref{eq:strain-hc}) and (\ref{eq:FTofEMRI}), the characteristic strain for inspiralling sources is given by \citep{FinnThorne}
\begin{equation}\label{eq:insphc}
h_\mathrm{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h_0 \;.
\end{equation}
Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals. Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.

@@

An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations:
\begin{equation}
h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
\end{equation}
For a linearly polarised GW, with $\tilde{h}(f)$ constant across the bandwidth $\Delta f$, this is approximately related to the characteristic strain via 
\begin{equation}
h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,,\label{eq:rela}
\end{equation}
where we have neglected the detector response functions (see section \ref{sec:principles}), which are of order unity. In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.
